---
title: "smires"
output: github_document
---

SMIRES is a COST Action addressing the Science and Management of Intermittent Rivers & Ephemeral Streams. SMIRES brings together more than 200 hydrologists, biogeochemists, ecologists, modellers, environmental economists, social researchers and stakeholders from 31 different countries to develop a research network for synthesising the fragmented, recent knowledge on IRES, improving our understanding of IRES and translating this into a science-based, sustainable management of river networks. More information about SMIRES can be found at <http://www.smires.eu/>. 

This git repository hosts the R-package `smires`, one of several outcomes of Working Group 1 (WG1, Prevalence, distribution and trends of IRES). Given time series of daily (weekly, monthly) discharges, its purpose is:


* to identify gauging stations that have an IRES flow regime;
* to calculate relevant flow and intermittency metrics.

# Installation of the R-package

In order to use the development version of the package `smires`  you will need to install the package `devtools` once. 

``` r
install.packages("devtools")
```


It provides the convenient commant `install_github()` which installs the most recent version of a package hosted on github. To do so, simply execute the following three lines:
``` r
library(devtools)
install_github("mundl/smires")
library(smires)
```

The current version is ``r packageVersion("smires")``. To load this package in an R session, `library(smires)` has to be called. 
``` r
library(smires)
packageVersion("smires")
```

# Examples

Each participating country was asked to suggest metrics and to submit a few time series with intermittent streamflow. As we plan to integrate the provided time series in the R package, we need permission of the copyright holder to make the data availabe. The column **include** of the following table indicates if the data can be published. 


### Time Series 

```{r results = 'asis', echo = FALSE} 

contrib_ts <- function(country, contact, ts = NA, publish = FALSE, comment = NA)
{
  data.frame("country" = country, "contact" = contact, "time series" = ts,
             "include" = if(is.na(ts)) NA else publish,
             "comment" = comment,
             check.names = FALSE, stringsAsFactors = FALSE)
}

series <- list(
  contrib_ts(country = "ch",
             contact = "ilja.vanmeerveld@geo.uzh.ch",
             ts = "Altlandenberg"),
  
  contrib_ts(country = "lt",
             contact = "hydro@mail.lei.lt"),
  
  contrib_ts(country = "it",
             contact = "giuseppe.verdiani@adb.puglia.it",
             ts = "Carapelle Torrent",
             comment = "Region Puglia"),
  
  contrib_ts(country = "uk",
             contact = "catsef@ceh.ac.uk",
             ts = "Balder at [Balderhead Reservoir](http://nrfa.ceh.ac.uk/data/station/info/25022), Ampney Brook at [Ampney St Peter](http://nrfa.ceh.ac.uk/data/station/info/39099)",
             publish = TRUE,
             comment = "Balder: human influence, Ampney Brook: chalk stream that dries naturally"), 
  
  contrib_ts(country = "pl",
             contact = "kazimierz_banasik@sggw.pl",
             ts = "Plachty Stare",
             comment = "uses hydrological year, starting with November"),  
  
  contrib_ts(country = "pl",
             contact = "rmrutkow@cyf-kr.edu.pl",
             ts = "Goryczkowa",
             comment = "only 3 years of observation"),
  
  contrib_ts(country = "fr", 
             contact = "eric.sauquet@irstea.fr",
             ts = "H1333010, H1503910, H1513210, H1603010, H1713010, H1932020"),
  
  contrib_ts(country = "es", 
             contact = "francesc.gallart@idaea.csic.es",
             ts = "Riu Manol"),
  
  contrib_ts(country = "it", 
             contact = "annamaria.degirolamo@ba.irsa.cnr.it",
             ts = "Celone, Salsola"),
  
  contrib_ts(country = "pt", 
             contact = "teresal@ipcb.pt",
             ts = "Coruche, Monforte, Pavia, Moinho"),  
  
  contrib_ts(country = "pt", 
             contact = "helena.alves@apambiente.pt",
             ts = "Sado"),

  contrib_ts(country = "gr", 
             contact = "rania.tzoraki@aegean.gr",
             ts = "Vrontamas"),

  contrib_ts(country = "sk", 
             contact = "kohnova@stuba.sk",
             ts = "Čierna voda, Bodva, Chlmec"))

series <- do.call(rbind, series)
series <- series[order(series$country), ]
row.names(series) <- NULL


options(knitr.kable.NA = '')
knitr::kable(series, caption = "Participants", format = "markdown")
```

<!--### Metrics -->


```{r results = 'asis', echo = FALSE} 

contrib_metric <- function(name, type = "NA", description = NA, comment = NA, 
                           dt = c("daily", "weekly", "monthly"), country)
{
  dt <- match.arg(dt)
  list("metric" = name, type = type, "description" = description,
       "\U0394t" = dt, "comment" = comment, country = sort(country))
}

metrics <- list(
  contrib_metric(name = "Number of days with zero flow",
                 description = "Total number of days with zero flow", 
                 type = "duration",
                 country = c("ch", "at", "gb")),
  contrib_metric(name = "Recession constant",
                 description = "Recession constant during the period with the lowest flow or slope of the flow duration curve", 
                 type = "magnitude",
                 country = c("ch", "es"))
)

countries <- unique(unlist(lapply(metrics, function(x) x$country), 
                           use.names = FALSE))
# matrix of country using a metric
cm <- used <- t(sapply(metrics, function(x) countries %in% x$country))
cm[used] <- "\U2714"
cm[!used] <- ""
colnames(cm) <- countries

metrics <- do.call(rbind, lapply(metrics, function(x) 
  as.data.frame(x[setdiff(names(x), "country")], as.is = TRUE) ))
metrics <- data.frame(metrics, cm)
metrics <- metrics[order(metrics$type), ]
row.names(metrics) <- NULL

#knitr::kable(metrics, caption = "Participants")
```




# Getting in Contact
In case you are interested or you want to contribute to the package `smires` (even though you are not part of the [SMIRES cost action](http://www.smires.eu/)) please contact <t.gauster@boku.ac.at>. 
