---
title: "Concept of the package"
author: "Tobias Gauster, Cath Sefton"
date: "`r Sys.Date()`"
output: 
  html_document:
    highlight: pygments
    number_sections: true
    toc: true
    toc_float: true
---

generated with `r R.version.string` and **smires** version `r packageVersion("smires")`


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```

This document was created with package version ``r packageVersion("smires")``.  Currently the following data sets are already included in the package:
```{r initialisation, warning=FALSE, message=FALSE}
library(smires)
data(package = "smires")$results[, "Item"]
```

Let's print the first observations of the river Balder at station Balderhead Reservoir:
```{r}
head(balder)
```

For the time being, each time series of discharges is a `data.frame` with the first column being the time index (of class `Date`) and the second column being the discharge. In this particular data set the unit of discharge is m³/s.
```{r}
library(ggplot2)
ggplot(balder, aes(time, discharge)) + geom_line() +
  labs(x = "", y = "discharge in m³/s")

```

A few checks are performed prior to the computation of indices. 

* time indices are unique
* time series is regular (daily, weekly or monthly)
* proportion of missing values
* proportion of values below or equal the measurement accuracy

```{r}
balder <- validate(balder, accuracy = 0.005)
```


# Units
For the time being only daily time series are supported. The default value of the threshold flow (0.001 m³/s) is only appropriate if your discharges are also in m³/s. If not, do not forget to choose a meaningful threshold. 

# Threshold 
A day is attributed to `no-flow` if the observed discharge is **smaller or equal** to the threshold discharge. Given a threshold of 0 m³/s and an observed discharge of 0 m³/s, this day is classified as `no-flow`. 

# Time intervals
```{r, message=FALSE, echo=FALSE}
head(find_spells(balder), 2)
```

A time interval like in the example above is constructed in a way that $duration = end - start$  implying that the end date is **not** part of the interval, it is a right opened interval: $[start,\ end)$. 