---
title: "smires -- Calculating Hydrological Metrics for Univariate Time Series"
author: "Tobias Gauster"
date: "2017-08-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(smires)
library(knitr)
```

## General definitions	

All indices are derived based on daily discharge hydrographs Q

No-flow spell:  period of time (at least one day) when discharge is less or equal some no-flow threshold (= detection limit): Q≤Q0 (default 1 l/s) 

No-flow event: a contiguous period of days with no-flow occurrence, terminated by rewetting period. Some pooling is usually applied to successive, dependent events in order to define an independent sequence of events.

Events are pooled unless the catchemnt has completely recovered from drought. 

All annual characteristic may be calculated based on calendar years or hydrological years (e.g. for European climates usually staring with Apr)

All annual characteristics can be calculated from the whole time series (all-season characteristics), or alternatively stratified by seasons (seasonal characteristics)




# Flow/No-flow spell characteristics

A characteristic (metric) that can be calculated for **no-flow** spells can also be calculated for **flow** spells. The following sections demonstrate just the calculation of no-flow spell characteristics. 


##  General characteristics
```{r, echo = FALSE, results = "asis"}
tbl <- metrics() %>% 
  filter(section == "General characteristics") %>%
  select(name, acronym, fun, description)
kable(tbl)
```


```{r}
no_flow_years(ampneyBrook)
```



##  Number of no-flow days
```{r, echo = FALSE, results = "asis"}
tbl <- metrics() %>% 
  filter(section == "Number of no-flow days") %>%
  select(name, acronym, fun, description)
kable(tbl)
```


```{r}
MAN(ampneyBrook)
```



## Duration
```{r, echo = FALSE, results = "asis"}
tbl <- metrics() %>% 
  filter(section == "Duration") %>%
  select(name, acronym, fun, description)
kable(tbl)
```


```{r}
MAMD(ampneyBrook)
```


## Timing and seasonality
```{r, echo = FALSE, results = "asis"}
tbl <- metrics() %>% 
  filter(section == "Timing and seasonality") %>%
  select(name, acronym, fun, description)
kable(tbl)
```


```{r}
Tau0(ampneyBrook)
```



# Rate of change before/after no-flow event
```{r, echo = FALSE, results = "asis"}
tbl <- metrics() %>% 
  filter(section == "Rate of change before/after no-flow event") %>%
  select(name, acronym, fun, description)
kable(tbl)
```


```{r}
k(ampneyBrook)
```

