---
title: "smires -- Calculating Hydrological Metrics for Univariate Time Series"
author: "Tobias Gauster"
date: "2017-08-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(knitr.kable.NA = '')
library(dplyr)
library(smires)
library(knitr)
library(lubridate)
```

# General definitions	

In our working group we agreed on the following definitions, most of them are already incorporated in the R package **smires**. 

* The terms characteristics, metrics and indices are used synonymously. 

* All indices are derived based on daily discharge hydrographs $Q$.  <span style="background-color: #D0D0FF;">The package itself just requires the times series to be regular (daily, weekly or monhtly). </span>

* No-flow *spell*:  period of time (at least one day) when discharge $Q$ is less or equal some no-flow threshold $Q_0$ (default 1 l/s, the detection limit): $Q ≤ Q_0$  

* No-flow *event*: a contiguous period of days with no-flow occurrence, terminated by rewetting period. Some pooling is usually applied to successive, dependent events in order to define an independent sequence of events. Events are pooled unless the catchemnt has completely recovered from drought. So events are pooled spells and events are independent of each other. <span style="background-color: #D0D0FF;">Pooling is not yet supported by the package. </span>

* As shown in the following figure, every characteristic can involve two aggregation levels the  *major* interval and the *minor* interval.  
In natural catchments most processes and quantities of the water cycle show an annual periodic behavior. The major interval is given by the hydrological understanding of a year. E.g. for European climates the hydrological year usually starts with April 1^st^. Alternatively calendar years starting on January 1^st^  can be used. A rule of thumb is that the major interval should be choosen such that events/spells are terminated before the end of a year.  
The minor interval allows a furhter partition of the year into eihter seasons, months or weeks. This is useful for calculating seasonal characteristics. If the minor interval is omitted, all-season characteristics are returned. 


Here, the function `smires` is used to demonstrate the meaning of major and minor intervals. The argument `major = 91` specifies that the major interval -- the hydrological year -- starts on the 91^st^ day of the year, which is April 1^st^. The default for minor intervals is to use months. The functions returns at table of spells, which we can ignore for the time being. The function is just used for the side effect of plotting the stream-flow permanence of the river 'Ampney Brook' at station 'Ampney St Peter' based on a threshold of 1 l/s.
```{r ampneyPerm, fig.width=10, fig.height=10}
ampney <- smires(ampneyBrook, major = 91, minor = intervals$month, rule = "onset", plot = TRUE)
```


The first record of the dataset exists for the date `r ampneyBrook$time[1]` which starts the `r smires:::.written_suffix(as.numeric(ampney$minor[1]))` month of our hydrological year `r ampney$major[1]`. The dataset covers in total `r (tot <- nlevels(ampney$major))` years, of which  `r (no <-  length(unique(ampney$major[ampney$state=="no-flow"])))` show a ``no-flow`` event. 


# Flow/No-flow spell characteristics

A characteristic  that can be calculated for **no-flow** spells can also be calculated for **flow** spells. The following sections demonstrate just the calculation of no-flow spell characteristics. 


##  General characteristics
Each of the following sections start with a table, showing all metrics attriubted to this certain section. This section 'General characteristics' comprises just one characteristic, the proportion of no-flow years  $f_0$ .
```{r, echo = FALSE, results = "asis"}
tbl <- metrics() %>% 
  filter(section == "General characteristics") %>%
  select(Name, Acronym, Function, Description)
kable(tbl)
```

Using the same example as before this metric is defined as 

$$f_0 = \frac{\textrm{Number of no-flow years}}{\textrm{Number of years in study period}} = \frac{`r no`}{`r tot`} = `r no/tot`= `r round(no/tot*100, 1)` \%$$

and can be calculated using the function:


```{r}
no_flow_years(ampneyBrook)
```



##  Number of no-flow days
```{r, echo = FALSE, results = "asis"}
tbl <- metrics() %>% 
  filter(section == "Number of no-flow days") %>%
  select(Name, Measure, Acronym, Function, Description)
kable(tbl)
```

Most IRES metrics are constructed in a similar way. To compute the mean annual number of no-flow days we have to do the following:

1. Given our daily discharge time series $Q$ we derive a **binary time series** of `flow` and `no-flow` states with respect to a theshold $Q_0$. This transforms our initial time series with `r nrow(ampneyBrook)` days of observations (rows) of a single variable (column `discharge`)  
```{r}
ampneyBrook
```
into a binary time series with `r nrow(ampney)` spells/events (rows) and `r ncol(ampney) -1` variables (columns). 
```{r}
ampney
```
Most of these newly created variables like `major`, `minor` and `group` are used for grouping. To compute mean annual number of no-flow days our main variable of interest is of course `duration`. 

2. As there can be easily several spells/events per year, for every year we need to aggregate the durations. To do so, in general we have to supply three arguments to the function `smires()`:
    + the variable we want to aggregate (in this case `duration` is implicitly used),
    + the interval we want to aggregate over (here hydrological years, `major = 91`)
    + and the function used to summarize all the durations (`fun_major = sum`).  
```{r}
smires(ampneyBrook, major = 91, fun_major = sum, complete = TRUE)
```
This yields us 


3. 


```{r}
MAN(ampneyBrook)
```



## Duration
```{r, echo = FALSE, results = "asis"}
tbl <- metrics() %>% 
  filter(section == "Duration") %>%
  select(Name, Measure, Acronym, Function, Description)
kable(tbl)
```


```{r}
MAMD(ampneyBrook)
```


## Timing and seasonality
```{r, echo = FALSE, results = "asis"}
tbl <- metrics() %>% 
  filter(section == "Timing and seasonality") %>%
  select(Name, Measure, Acronym, Function, Description)
kable(tbl)
```


```{r}
tau0(ampneyBrook)
```



# Rate of change before/after no-flow event
```{r, echo = FALSE, results = "asis"}
tbl <- metrics() %>% 
  filter(section == "Rate of change before/after no-flow event") %>%
  select(Name, Measure, Acronym, Function, Description)
kable(tbl)
```


```{r}
k(ampneyBrook)
```

