---
title: "Concept of the package"
author: "Tobias Gauster, Cath Sefton"
date: "February 10, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```

This document was created with package version ``r packageVersion("smires")``.  Currently the following datasets are already included in the package:
```{r initialisation, warning=FALSE, message=FALSE}
library(smires)
data(package = "smires")$results[, "Item"]
```

Let's print the first observations of the river Balder at station Balderhead Reservoir:
```{r}
head(balder)
```

For the time being, each time series of discharges is a `data.frame` with the first column being the time index (of class `Date`) and the second column being the discharge. In this particular dataset the unit of discharge is m³/s.
```{r}
library(ggplot2)
ggplot(balder, aes(time, discharge)) + geom_line() +
  labs(x = "", y = "discharge in m³/s")

```

A few checks are performed prior to the computation of indices. 

* time indices are unique
* time series is regular (daily, weekly or monthly)
* proportion of missing values
* proportion of values below or equal the measurement accurancy

```{r}
balder <- check_ts(balder, accuracy = 0.005)
```
